---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# touRdemols

<!-- badges: start -->
<!-- badges: end -->

The package touRdemols is a collection of functions that can be used to efficiently calculate and display the standings in a Tour de Mols competition. 

## Background

The competition is a central part of the Arthropod Field Course for first year undergraduate students at Aarhus University, but can probably be applied anywhere in the world for any larger taxon. The competition dates back to ...

## The practical stuff

Multiple teams compete to find the most species. Teams consist of varying numbers of persons, but the points are standardized to a team size of four.

The point system rewards a number of things:  
  
* Encountering a new species (1 point)  
* Encountering a new family (defaults to 1 point*)  
* Encountering a new order (defaults to 3 points*)  
* Encountering a new class (defaults to 5 points*)  
* Getting good at a family. When you encounter a certain number of species within a family an extra point is rewarded per species (threshold is four species*). This is used to encourage people to get good a specific key.  
* Encountering red listed species. Species listed in the categories NT, VU, EN, CR, RE on the regional IUCN Redlist are rewarded with more points than other species.*  

'*' values can be set differently with the parameters in calculate_standings().  

## Installation

You can install the released version of touRdemols from [GitHub](https://github.com/) with:

``` {r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("andreasbennetsenboe/touRdemols")
```

## Example

The folder 'exampledata' contains seven data sets from teams that competed against each other in the Tour de Mols. We use these here to illustrate how to use the package. Furthermore it contains a little test data set to show how the data should be structured:

```{r, echo = FALSE}
utils::head(read.csv2("exampledata/test.csv"))
```

For people that are not used to working with R or .csv files one can use the 'pointscheme.xlsx' to record your observations and then save as a .csv file and just name it the integer that represents your group. When all schemes are sent to you at the end of the day you can start importing. Do this as a list of teams:
 
```{r example}
library(touRdemols)

list.of.teams <- list(
  team1 <- read.csv2("exampledata/1.csv", sep = ";"),
  team2 <- read.csv2("exampledata/2.csv", sep = ";"), 
  team3 <- read.csv2("exampledata/3.csv", sep = ";"), 
  team4 <- read.csv2("exampledata/4.csv", sep = ";"), 
  team5 <- read.csv2("exampledata/5.csv", sep = ","), 
  team6 <- read.csv2("exampledata/6.csv", sep = ","), 
  team7 <- read.csv2("exampledata/7.csv", sep = ";")
)
```

You then feed this vector to the function calculate_standings(). If you want to change anything in the point system, the function has a number of parameters for this, see the help file ?calculate_standings for details.

```{r, echo = FALSE}
#calculate_standings(teams = list.of.teams,
#                    instructors = c("ABB", "CK", "SK", "CB")
#                    )
```


